---
title: "Neurosurgery survey"
author: "Pavel Dusek"
date: "05 04 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
```

## Software Info
Built with R version `r getRversion()`, dplyr version `r packageVersion("dplyr")`, tidyr version `r packageVersion("tidyr")`, ggplot2 version `r packageVersion("ggplot2")`.

## Loading and Cleaning Data 
We load the dataset:
```{r}
df <- read.csv("palliative_neurosurgery_survey.csv")
```

Order factors
```{r}
df <- df %>% mutate(
 Demographics_On_Call = factor(Demographics_On_Call, levels = c( "<3 days a month", "3-5 days a month", "5-10 days a month ", ">10 days a month")),
 Demographics_years_of_practise = factor(Demographics_years_of_practise, levels = c( "<5", "5-10", "10-20", ">20" )),
 Futile_Surgery = factor(Futile_Surgery, levels = c( "Never", "Rarely (once a year)", "Sometimes (2-5 a year)", "Often (>5 a year)" )),
 Burnout = factor(Burnout, levels = c( "NO", "MAYBE", "YES" )),
 Moral_Distress_Experience = factor(Moral_Distress_Experience, levels = c( "Strongly disagree", "Generally disagree", "Generally agree", "Strongly agree" )),
 Leaving_Position = factor(Leaving_Position, levels = c( "No, I have never considered leaving or left a position.", "Yes, I considered leaving but did not leave.", "Yes, I left a position.")),
 Communication_Training_Comfortability = factor( Communication_Training_Comfortability, levels = c( "None at all", "A little", "A moderate amount", "A lot", "A great deal" ) )
)
```

Custom function to print out descriptive statistics:
```{r}
describe <- function( df, column_names, caption ) {
 df %>%
 as_tibble() %>%
 select( all_of(column_names) ) %>%
 pivot_longer( cols = all_of(column_names) ) %>%
 group_by(name) %>%
 summarize_at(
  .vars = vars(value),
  .funs = list(
   min = min,
   Q1 = ~quantile(., probs = 0.25),
   median = median,
   Q3 = ~quantile(., probs = 0.75),
   max= max,
   mean = ~round(mean(.), digits = 2),
   sd = ~round(sd(.), digits = 2)
  )
 ) %>%
 kable( caption = caption )
}

frequency_table <- function( df, variable ) {
 df %>%
 select_at( variable ) %>%
 group_by_at( variable ) %>%
 count() %>%
 mutate( perc = round( (100 * n )/dim(df)[1], digits = 2) ) %>%
 kable( caption = paste( variable, "Counts" ) )
}

describe_univariate <- function( df, variable ) {
 df %>%
 select( all_of( c(variable, "Moral_Distress_Score") ) ) %>%
 group_by_at(variable) %>%
 summarize_at(
  vars(Moral_Distress_Score),
  .funs = list(
   min = min,
   Q1 = ~quantile(., probs = 0.25),
   median = median,
   Q3 = ~quantile(., probs=0.75),
   max= max,
   mean = ~round(mean(.), digits=2),
   sd = ~round(sd(.), digits=2)
  )
 ) %>%
 arrange( desc( median ) ) %>%
 kable( caption = paste("Moral Distress grouped by", variable ) )
}

plot_univariate <- function( df, variable ) {
 ggplot( df, aes( x = .data[[variable]], y = Moral_Distress_Score) ) +
 geom_violin(trim = FALSE) +
 geom_boxplot( width = 0.1 ) +
 ggtitle( paste( "Moral Distress Score by", variable ) )
}
```

## Oldenburg Burnout Inventory (OLBI)
```{r}
describe( df = df, column_names = c("OLBI", "OLBI.D", "OLBI.E"), caption = "OLBI Descriptive Statistics" )
```

## Moral Distress Description
### Total Score and Subscores
```{r}
ggplot( df, aes( x = Moral_Distress_Score ) ) +
 geom_histogram( bins = 20 ) +
 ggtitle( "Histogram of Moral_Distress_Score" )
```

```{r}
describe(
	df = df,
	column_names = c("Moral_Distress_Score", "Moral_Distress_Score_frequency", "Moral_Distress_Score_intensity"), 
	caption = "Moral Distress Score Descriptive Statistics"
)
```
### Each Question Frequency
```{r}
frequency_columns <- c(
	"Moral_Distress_1_frequency",
	"Moral_Distress_2_frequency",
	"Moral_Distress_3_frequency",
	"Moral_Distress_4_frequency",
	"Moral_Distress_5_frequency",
	"Moral_Distress_6_frequency",
	"Moral_Distress_7_frequency",
	"Moral_Distress_8_frequency"
)
describe(
	 df = df,
	 column_names = frequency_columns,
	 caption = "Moral Distress Frequency Questions Descriptive Statistics"
)
```
Questions sorted by mean frequency (highest to lowest):
```{r}
df %>% as_tibble() %>% select( all_of( frequency_columns ) ) %>% pivot_longer( all_of(frequency_columns) ) %>% group_by( name ) %>% summarize( mean = mean(value) ) %>% arrange( desc(mean) ) %>% select( name )
```

### Each Question Intensity
```{r}
intensity_columns <- c(
	"Moral_Distress_1_intensity",
	"Moral_Distress_2_intensity",
	"Moral_Distress_3_intensity",
	"Moral_Distress_4_intensity",
	"Moral_Distress_5_intensity",
	"Moral_Distress_6_intensity",
	"Moral_Distress_7_intensity",
	"Moral_Distress_8_intensity"
)
describe(
	 df = df,
	 column_names = intensity_columns,
	 caption = "Moral Distress Intensity Questions Descriptive Statistics"
)
```
Questions sorted by mean intensity (highest to lowest):
```{r}
df %>% as_tibble() %>% select( all_of( intensity_columns ) ) %>% pivot_longer( all_of(intensity_columns) ) %>% group_by( name ) %>% summarize( mean = mean(value) ) %>% arrange( desc(mean) ) %>% select( name )
```

## Moral Distress Univariate Analysis
### Gender
```{r}
frequency_table( df, "Demographics_gender" )
describe_univariate( df, "Demographics_gender" )
plot_univariate( df, "Demographics_gender" )
wilcox.test(
	Moral_Distress_Score ~ Demographics_gender,
	data = df %>%
		select( Moral_Distress_Score, Demographics_gender ) %>%
		filter( Demographics_gender != "Decline to answer")
)
```

### Years of practice
```{r}
frequency_table( df, "Demographics_years_of_practise" )
describe_univariate( df, "Demographics_years_of_practise" )
plot_univariate( df, "Demographics_years_of_practise" )
kruskal.test(
	Moral_Distress_Score ~ Demographics_years_of_practise,
	data = df %>% filter( Demographics_years_of_practise != "Decline to answer" )
)
```

### Position
```{r}
frequency_table( df, "Demographics_position" )
describe_univariate( df, "Demographics_position" )
plot_univariate( df, "Demographics_position" )
wilcox.test(
	Moral_Distress_Score ~ Demographics_position,
	data = df
)
```

### Trauma Center
```{r}
frequency_table( df, "Demographics_Trauma_Center" )
describe_univariate( df, "Demographics_Trauma_Center" )
plot_univariate( df, "Demographics_Trauma_Center" )
wilcox.test(
    Moral_Distress_Score ~ Demographics_Trauma_Center,
    data = df %>% filter( Demographics_Trauma_Center != "Decline to answer" )
)
```
)

### On Call
```{r}
frequency_table( df, "Demographics_On_Call" )
describe_univariate( df, "Demographics_On_Call" )
plot_univariate( df, "Demographics_On_Call" )
kruskal.test(
    Moral_Distress_Score ~ Demographics_On_Call,
    data = df %>% filter( Demographics_On_Call != "Decline to answer" )
)
```

### Futile Surgery Frequency
```{r}
frequency_table( df, "Futile_Surgery" )
describe_univariate( df, "Futile_Surgery" )
plot_univariate( df, "Futile_Surgery" )
  kruskal.test(
    Moral_Distress_Score ~ Futile_Surgery,
    data = df
  )
```

### Burnout self-assessment
```{r}
frequency_table( df, "Burnout" )
describe_univariate( df, "Burnout" )
plot_univariate( df, "Burnout" )
  kruskal.test(
    Moral_Distress_Score ~ Burnout,
    data = df
  )
```

### Experienced moral distress during last year
```{r}
frequency_table( df, "Moral_Distress_Experience" )
describe_univariate( df, "Moral_Distress_Experience" )
plot_univariate( df, "Moral_Distress_Experience" )
kruskal.test(
	Moral_Distress_Score ~ Moral_Distress_Experience,
	data = df
)
```

### Leaving a position because of moral distress
```{r}
frequency_table( df, "Leaving_Position" )
describe_univariate( df, "Leaving_Position" )
plot_univariate( df, "Leaving_Position" )
kruskal.test(
	Moral_Distress_Score ~ Leaving_Position,
	data = df
)
```

## Multivariate Regression Analysis of Moral Distress Score
```{r}
model <- lm(
   Moral_Distress_Score ~ Demographics_years_of_practise + Demographics_Trauma_Center + Futile_Surgery + Burnout + 1,
   data = df
)
summary(model)
```

```{r}
model <- lm(
   Moral_Distress_Score ~ Demographics_years_of_practise + Demographics_Trauma_Center + Futile_Surgery + Burnout + PHQ9 + 1,
   data = df
)
summary(model)
```

## Communication training
* Communication_Training_1: Did you seek out any communication training to help you with hyperintense communication during you residency?  (YES, NO)
* Communication_Training_2: Did you receive training specific to lead goal of care communication during your residency?  (YES, NO)
Communication_Training_3: Did you seek out any communication training to help you with hyperintense communication after you graduated?  (YES, NO)
* Communication_Training_4: Would you be interested in communication training to help you navigate hyperintense communication? (YES, NO)
* Communication_Training_Skills: For the following statement, please indicate how much you agree or disagree: I have the necessary skills to discuss end of life goals in neurosurgery patients. (Strongly disagree, Generally disagree, Generally agree, Strongly agree)

```{r}
df %>%
 select( Communication_Training_1, Communication_Training_2, Communication_Training_3, Communication_Training_4 ) %>%
 pivot_longer(c(Communication_Training_1, Communication_Training_2, Communication_Training_3, Communication_Training_4)) %>%
 table() %>%
 as_tibble() %>%
 pivot_wider(names_from = value, values_from = n) %>%
 mutate( NO_perc = round( (100 * NO)/dim(df)[1], digits = 2), YES_perc = round( (100 * YES)/dim(df)[1], digits = 2) ) %>%
 kable(caption = "Communication Training Summary Statistics")
```

Communication_Training_Skills: For the following statement, please indicate how much you agree or disagree: I have the necessary skills to discuss end of life goals in neurosurgery patients.
```{r}
df %>%
 select( Communication_Training_Skills ) %>%
 table() %>%
 as_tibble() %>%
 mutate( Perc = round( (100 * n) / dim(df)[1], digits = 2) ) %>%
 kable(caption = "Communication Training Skills Summary Statistics")
```

Communication_Training_Comfortability: How comfortable do you feel engaging in hyper-intense conversations. Hyper-intense conversations.are defined as situations in which you have to break bad news to a patient or patient’s family member (death, poor prognosis, or withdrawal of life sustaining treatments).
```{r}
df %>%
 select( Communication_Training_Comfortability ) %>%
 table() %>%
 as_tibble() %>%
 mutate( Perc = round( (100 * n )/dim(df)[1] , digits = 2) ) %>%
 kable( caption = "Communication Training Comfortability" )
```

## Palliative Care
* Palliative_Care_Availability: Is palliative care available as a specialty service at your institution?  (YES, NO)
* Palliative_Care_TBI: On which patient do you routinely consult palliative care? (check all that apply) Severe TBI (YES, NO)
* Palliative_Care_GBM: On which patient do you routinely consult palliative care? (check all that apply) Recurrent GBM (YES, NO)
* Palliative_Care_Metastatic: On which patient do you routinely consult palliative care? (check all that apply) Widely metastatic disease (YES, NO)
* Palliative_Care_Rarely: On which patient do you routinely consult palliative care? (check all that apply) I rarely consult palliative care (YES, NO)
* Palliative_Care_Other: On which patient do you routinely consult palliative care? (check all that apply) x Other...  (NO, free text)
* Palliative_Care_Reason1: What are the reasons you consult palliative care? (check all that apply) Patient likely to die during this admission (YES, NO)
* Palliative_Care_Reason2: What are the reasons you consult palliative care? (check all that apply) Difficult family dynamics (YES, NO)
* Palliative_Care_Reason3: What are the reasons you consult palliative care? (check all that apply) End of life discussion is needed (YES, NO)
* Palliative_Care_Reason4: What are the reasons you consult palliative care? (check all that apply) Patient’s/family’s request (YES, NO)
* Palliative_Care_Reason5: What are the reasons you consult palliative care? (check all that apply) Time constraints in hyperintense conversations (YES, NO)
* Palliative_Care_Reason6: What are the reasons you consult palliative care? (check all that apply) Lack of training in having hyperintense conversations (YES, NO)
* Palliative_Care_Reason7: What are the reasons you consult palliative care? (check all that apply) Other (NO, free text)

```{r}
df %>%
 select( 
  Palliative_Care_Availability,
  Palliative_Care_TBI,
  Palliative_Care_GBM,
  Palliative_Care_Metastatic,
  Palliative_Care_Rarely,
  Palliative_Care_Reason1,
  Palliative_Care_Reason2,
  Palliative_Care_Reason3,
  Palliative_Care_Reason4,
  Palliative_Care_Reason5,
  Palliative_Care_Reason6
 ) %>%
 pivot_longer( c(
  Palliative_Care_Availability,
  Palliative_Care_TBI,
  Palliative_Care_GBM,
  Palliative_Care_Metastatic,
  Palliative_Care_Rarely,
  Palliative_Care_Reason1,
  Palliative_Care_Reason2,
  Palliative_Care_Reason3,
  Palliative_Care_Reason4,
  Palliative_Care_Reason5,
  Palliative_Care_Reason6
 )) %>%
 table() %>%
 as_tibble() %>%
 pivot_wider(names_from = value, values_from = n) %>%
 mutate( NO_perc = round( (100 * NO)/dim(df)[1], digits = 2), YES_perc = round( (100 * YES)/dim(df)[1], digits = 2) ) %>%
 kable(caption = "Communication Training Summary Statistics")
```

Palliative_Care_Other: On which patient do you routinely consult palliative care? (check all that apply) x Other...  (NO, free text)
```{r}
df %>% select( Palliative_Care_Other ) %>% table() %>% as_tibble()
```

Palliative_Care_Reason7: What are the reasons you consult palliative care? (check all that apply) Other (NO, free text)
```{r}
df %>% select( Palliative_Care_Reason7 ) %>% table() %>% as_tibble()
```

## Depression
```{r}
describe( df = df, column_names = c("PHQ9"), caption = "PHQ9 Descriptive Statistics" )
```

```{r}
ggplot( df, aes( x = PHQ9 ) ) +
 geom_histogram( bins = 15 ) +
 ggtitle( "Histogram of PHQ9" )
```

### Depression and Moral Distress
Scatter
```{r}
ggplot( df, aes( x = Moral_Distress_Score, y = PHQ9 ) ) +
 geom_point() +
 geom_smooth(method = "lm" , formula = "y ~ x") +
 ggtitle( "Scatter plot PHQ9 vs. Moral Distress Score" )
```

Pearson's correlation coefficient of Moral Distress Score and PHQ9:
```{r}
cor.test(df$PHQ9, df$Moral_Distress_Score)
```

Linear regression model of Moral Distress Score and PHQ9:
```{r}
model <- lm ( PHQ9 ~ Moral_Distress_Score, df)
summary(model)
```

### Depression and Burnout
```{r}
ggplot( df, aes( x = OLBI, y = PHQ9 ) ) +
 geom_point() +
 geom_smooth(method = "lm" , formula = "y ~ x") +
 ggtitle( "Scatter plot PHQ9 vs. Oldenburg Burnout Inventory" )
```

Pearson's correlation coefficient of Burnout and PHQ9:
```{r}
cor.test(df$PHQ9, df$OLBI)
```

Linear regression model of OLBI and PHQ9:
```{r}
model <- lm ( PHQ9 ~ OLBI, df)
summary(model)
```

Self-reported Burnout:
```{r}
kruskal.test(
	PHQ9 ~ Burnout,
	data = df
)
```

```{r}
ggplot( df, aes( x = Burnout, y = PHQ9 ) ) +
 geom_violin(trim = FALSE) +
 geom_boxplot( width = 0.1 ) +
 ggtitle( "PHQ9 by self-reported Burnout" )
```

### Depression Multivariate Regression Analysis
```{r}
model <- lm ( PHQ9 ~ Moral_Distress_Score + OLBI, df)
summary(model)
```

```{r}
ggplot( df, aes( x = Moral_Distress_Score, y = OLBI, size = PHQ9 ) ) +
 geom_point() +
 ggtitle("PHQ9 predicted by Moral_Distress_Score and OLBI")
```
