---
title: "Neurosurgery survey"
author: "Pavel Dusek"
date: "02 04 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
```

## Software Info
Built with R version `r getRversion()`, dplyr version `r packageVersion("dplyr")`, tidyr version `r packageVersion("tidyr")`.

## Data Description
We load the dataset:
```{r}
df <- read.csv("palliative_neurosurgery_survey.csv")
df <- df %>% rename( Futile_Surgery = Moral_Distress_frequency, Leaving_Position = Moral_Distress_Futile_surgery )
```

Order factors
```{r}
df$Demographics_On_Call <- factor(
 df$Demographics_On_Call,
 levels = c(
  "<3 days a month",
  "3-5 days a month",
  "5-10 days a month ",
  ">10 days a month"
 )
)

df$Demographics_years_of_practise <- factor(
 df$Demographics_years_of_practise,
 levels = c( "<5", "5-10", "10-20", ">20" )
)

df$Futile_Surgery <- factor(
 df$Futile_Surgery,
 levels = c( "Never", "Rarely (once a year)", "Sometimes (2-5 a year)", "Often (>5 a year)" )
)

df$Burnout <- factor(
 df$Burnout,
 levels = c( "NO", "MAYBE", "YES" )
)

df$Moral_Distress_Experience <- factor(
 df$Moral_Distress_Experience,
 levels = c( "Strongly disagree", "Generally disagree", "Generally agree", "Strongly agree" )
)

df$Leaving_Position <- factor(
 df$Leaving_Position,
 levels = c(
  "No, I have never considered leaving or left a position.",
  "Yes, I considered leaving but did not leave.",
  "Yes, I left a position."
 )
)
```

Custom function to print out descriptive statistics:
```{r}
describe <- function( df, column_names, caption ) {
 df %>%
 as_tibble() %>%
 select( all_of(column_names) ) %>%
 pivot_longer( cols = all_of(column_names) ) %>%
 group_by(name) %>%
 summarize_at(
  .vars = vars(value),
  .funs = list(
   min = min,
   Q1 = ~quantile(., probs = 0.25),
   median = median,
   Q3 = ~quantile(., probs = 0.75),
   max= max,
   mean = ~round(mean(.), digits = 2),
   sd = ~round(sd(.), digits = 2)
  )
 ) %>%
 kable( caption = caption )
}

frequency_table <- function( df, variable ) {
 df %>%
 select_at( variable ) %>%
 group_by_at( variable ) %>%
 count() %>%
 kable( caption = paste( variable, "Counts" ) )
}

describe_univariate <- function( df, variable ) {
 df %>%
 select( all_of( c(variable, "Moral_Distress_Score") ) ) %>%
 group_by_at(variable) %>%
 summarize_at(
  vars(Moral_Distress_Score),
  .funs = list(
   min = min,
   Q1 = ~quantile(., probs = 0.25),
   median = median,
   Q3 = ~quantile(., probs=0.75),
   max= max,
   mean = ~round(mean(.), digits=2),
   sd = ~round(sd(.), digits=2)
  )
 ) %>%
 arrange( desc( median ) ) %>%
 kable( caption = paste("Moral Distress grouped by", variable ) )
}

plot_univariate <- function( df, variable ) {
 ggplot( df, aes( x = .data[[variable]], y = Moral_Distress_Score) ) +
 geom_violin(trim = FALSE) +
 geom_boxplot( width = 0.1 ) +
 ggtitle( paste( "Moral Distress Score by", variable ) )
}
```

## Oldenburg Burnout Inventory (OLBI)
```{r}
describe( df = df, column_names = c("OLBI", "OLBI.D", "OLBI.E"), caption = "OLBI Descriptive Statistics" )
```

## Moral Distress Description
### Total Score and Subscores
```{r}
describe(
	df = df,
	column_names = c("Moral_Distress_Score", "Moral_Distress_Score_frequency", "Moral_Distress_Score_intensity"), 
	caption = "Moral Distress Score Descriptive Statistics"
)
```
### Each Question Frequency
```{r}
frequency_columns <- c(
	"Moral_Distress_1_frequency",
	"Moral_Distress_2_frequency",
	"Moral_Distress_3_frequency",
	"Moral_Distress_4_frequency",
	"Moral_Distress_5_frequency",
	"Moral_Distress_6_frequency",
	"Moral_Distress_7_frequency",
	"Moral_Distress_8_frequency"
)
describe(
	 df = df,
	 column_names = frequency_columns,
	 caption = "Moral Distress Frequency Questions Descriptive Statistics"
)
```
Questions sorted by mean frequency (highest to lowest):
```{r}
df %>% as_tibble() %>% select( all_of( frequency_columns ) ) %>% pivot_longer( all_of(frequency_columns) ) %>% group_by( name ) %>% summarize( mean = mean(value) ) %>% arrange( desc(mean) ) %>% select( name )
```

### Each Question Intensity
```{r}
intensity_columns <- c(
	"Moral_Distress_1_intensity",
	"Moral_Distress_2_intensity",
	"Moral_Distress_3_intensity",
	"Moral_Distress_4_intensity",
	"Moral_Distress_5_intensity",
	"Moral_Distress_6_intensity",
	"Moral_Distress_7_intensity",
	"Moral_Distress_8_intensity"
)
describe(
	 df = df,
	 column_names = intensity_columns,
	 caption = "Moral Distress Intensity Questions Descriptive Statistics"
)
```
Questions sorted by mean intensity (highest to lowest):
```{r}
df %>% as_tibble() %>% select( all_of( intensity_columns ) ) %>% pivot_longer( all_of(intensity_columns) ) %>% group_by( name ) %>% summarize( mean = mean(value) ) %>% arrange( desc(mean) ) %>% select( name )
```

## Moral Distress Univariate Analysis
### Gender
```{r}
frequency_table( df, "Demographics_gender" )
describe_univariate( df, "Demographics_gender" )
plot_univariate( df, "Demographics_gender" )
wilcox.test(
	Moral_Distress_Score ~ Demographics_gender,
	data = df %>%
		select( Moral_Distress_Score, Demographics_gender ) %>%
		filter( Demographics_gender != "Decline to answer")
)
```

### Years of practice
```{r}
frequency_table( df, "Demographics_years_of_practise" )
describe_univariate( df, "Demographics_years_of_practise" )
plot_univariate( df, "Demographics_years_of_practise" )
kruskal.test(
	Moral_Distress_Score ~ Demographics_years_of_practise,
	data = df %>% filter( Demographics_years_of_practise != "Decline to answer" )
)
```

### Position
```{r}
frequency_table( df, "Demographics_position" )
describe_univariate( df, "Demographics_position" )
plot_univariate( df, "Demographics_position" )
wilcox.test(
	Moral_Distress_Score ~ Demographics_position,
	data = df
)
```

### Trauma Center
```{r}
frequency_table( df, "Demographics_Trauma_Center" )
describe_univariate( df, "Demographics_Trauma_Center" )
plot_univariate( df, "Demographics_Trauma_Center" )
wilcox.test(
    Moral_Distress_Score ~ Demographics_Trauma_Center,
    data = df %>% filter( Demographics_Trauma_Center != "Decline to answer" )
)
```
)

### On Call
```{r}
frequency_table( df, "Demographics_On_Call" )
describe_univariate( df, "Demographics_On_Call" )
plot_univariate( df, "Demographics_On_Call" )
kruskal.test(
    Moral_Distress_Score ~ Demographics_On_Call,
    data = df %>% filter( Demographics_On_Call != "Decline to answer" )
)
```

### Futile Surgery Frequency
```{r}
frequency_table( df, "Futile_Surgery" )
describe_univariate( df, "Futile_Surgery" )
plot_univariate( df, "Futile_Surgery" )
  kruskal.test(
    Moral_Distress_Score ~ Futile_Surgery,
    data = df
  )
```

### Burnout self-assessment
```{r}
frequency_table( df, "Burnout" )
describe_univariate( df, "Burnout" )
plot_univariate( df, "Burnout" )
  kruskal.test(
    Moral_Distress_Score ~ Burnout,
    data = df
  )
```

### Experienced moral distress during last year
```{r}
frequency_table( df, "Moral_Distress_Experience" )
describe_univariate( df, "Moral_Distress_Experience" )
plot_univariate( df, "Moral_Distress_Experience" )
kruskal.test(
	Moral_Distress_Score ~ Moral_Distress_Experience,
	data = df
)
```

### Leaving a position because of moral distress
```{r}
frequency_table( df, "Leaving_Position" )
describe_univariate( df, "Leaving_Position" )
plot_univariate( df, "Leaving_Position" )
kruskal.test(
	Moral_Distress_Score ~ Leaving_Position,
	data = df
)
```

## Multivariate Regression Analysis of Moral Distress Score

## Communication training

## Palliative Care

## Depression

